---
title: "100 top tracks of 2018 From Spotify"
author: "Tingting Zheng"
date: "February 23, 2020"
output:
  html_document:
    toc: yes
    toc_depth: '2'
    df_print: paged
  pdf_document:
    toc: yes
    toc_depth: '2'
---


# Project Description

Nowadays we use Spotify almost every day not only for entertainment, but also for study. We are going to to look for patterns in the audio features of the songs. We will try to predict one audio feature based on the others. 


## Research Question


**Question: Will the loudness affect the energy of a song?**

...


## Variables

*The explanatory variable is the loudness. The response variable is the energy.*



# Exploratory Data Analysis (EDA)

# Data Description

*This is a data-set of 100 top tracks of 2018 From Spotify. Each row represents a track. The id column is the categorical variable which is the Spotify’s identification of the song. The name column is also the categorical variable and shows the name of each video. The artists column is the categorical variable which displays artist(s) of the song. The danceability column is a numerical variable and displays how suitable a track is for dancing based on a combination of musical elements including tempo, rhythm stability, beat strength, and overall regularity. The energy column is the numerical variable which is a measure from 0.0 to 1.0 and represents a perceptual measure of intensity and activity. The key column is the categorical variable. The key the track is in. Integers map to pitches using standard pitch Class notation. E.g. 0=C, 2=D, and so on. The loudness column is the numerical variable. The overall loudness of a track in decibels (dB). The mode column is the quantitative variable. Mode indicates the modality (major or minor) of a track, the type of scale from which its melodic content is derived. The speechiness column is the quantitative variable. Speechiness detects the presence of spoken words in a track. The more exclusively speech-like the recording (e.g. talk show, audio book, poetry), the closer to 1.0 the attribute value. The acousticness column is the quantitative variable. A confidence measure from 0.0 to 1.0 of whether the track is acoustic. The instrumentalness column is the quantitative variable. Predicts whether a track contains no vocals. The liveness column is the quantitative variable which detects the presence of an audience in the recording. The valence column is the quantitative variable. It’s a measure from 0.0 to 1.0 describing the musical positiveness conveyed by a track. The tempo column is the quantitative variable. The overall estimated tempo of a track in beats per minute (BPM). The duration_ms column is the quantitative variable. It is the duration of the track in milliseconds. The time_signature column is the quantitative variable which is an estimated overall time signature of a track. The time signature (meter) is a notational convention to specify how many beats are in each bar (or measure). The explanatory variable is the loudness. The response variable is the energy.*



# Statistical Analysis 

*Read the data into R and conduct some initial summaries of your exploratory and response data.*
```{r}
top2018<-read.csv("top2018.csv")
attach(top2018) # The database "top2018 is attached to the R search path
summary(top2018) # To produce result summaries of the data from top2018.
dim(top2018) # to get dimensions Of data top2018
qqnorm(top2018$energy, pch = 1, frame = FALSE, main = "Normal Q-Q Plot for Energy") # produces a normal QQ plot of the value energy
qqline(top2018$energy, col = "steelblue", lwd = 2) # adds a line to a “theoretical”, by default normal, quantile-quantile plot
qqnorm(top2018$loudness, pch = 1, frame = FALSE, main = "Normal Q-Q Plot for Loudness") # produces a normal QQ plot of the value loudness
qqline(top2018$loudness, col = "steelblue", lwd = 2) # adds a line to a “theoretical”, by default normal, quantile-quantile plot
hist(top2018$energy, main = "histogram of engergy", xlab = "energy", ylab = "frequency") # produce a histogram of energy
hist(top2018$loudness, main = "histogram of loudness", xlab = "loudness", ylab = "dB") # produce a histogram of loudness
boxplot(top2018$energy, main = "boxplot of engergy", xlab = "energy", ylab = "frequency") # produce a boxplot of engergy
boxplot(top2018$loudness, main = "boxplot of loudness", xlab = "loudness", ylab = "dB") # produce a boxplot of loudness
```
*From the R code of the dimension of the dataset, there are total 100 data we have.*
*From the summery code, we can conclude the range of the energy is from 0.2960 to 0.9090. The range of the loudness is from -10.109 to -2.384.*
*From the normal QQ plots and histograms for Loudness and Energy, we can say that for both Energy and loudness, they are both normal distributed.*
*When we look at the box plots of energy and loudness, we will find that there is no obvious outlier in this dataset.*

*Create a scatter plot of your explanatory and response variables. Describe the relationship’s form, direction and strength and any outliers or features.*
```{r}
plot(x = top2018$loudness, 
     y = top2018$energy, 
     main = "Energy vs Loudness",
     xlab = "Loudness", 
     ylab = "Energy",  
     pch = 16) # A scatterplot of energy and loudness. With loudness in the x-axis, energy in the y-axis.

```
*From the scatterplot, we can find that there is a positive, moderate, and linear relationship between Variable Loudness and Energy. And there is no obvious outlier shown in the scatter plot.*
*Use R to calculate the correlation coefficient and its 99% confidence interval.*
```{r}
cor.test(loudness,energy,conf.level=0.99) # Test For Correlation Between loudness and energy and it's 99% confidence interval
```
*The 99% confidence interval is from 0.5869744 to 0.8324640. The correlation coefficient is 0.7327191.*

*Use R to create a linear model between two variables and save it to a variable called fit1.*
```{r}
Fit1=lm(energy~loudness,data=top2018) #creat a linear model between energy and loudness
summary(Fit1) # To produce result summaries of the linear model Fit1
```
*y=0.05980x+0.99857, where y is the energy, x is the loudness, b1=0.05980, b0=0.99857.*
*For the regression model, it indicates that when loudness equals to 0, the energy is 0.99857. And for each unit increased in loudness, the energy will increase 0.05980.*

*identify the R2 and adjusted R2 values*
```{r}
summary(Fit1) # To produce result summaries of the linear model Fit1
```
*R2 equals to 0.5369(53.69%), and adjusted R2 equals to 0.5322(53.22%).*
*Add the line of best fit to the scatter plot.*
```{r}
plot(x = top2018$loudness, 
     y = top2018$energy,  
     main = "Energy vs Loudness",
     xlab = "Loudness", 
     ylab = "Energy", 
     pch = 16) # produce the scatter plot of energy and loudness, with loudness in the x-axis, energy in the y-axis
abline(lm(energy~loudness, data=top2018), col='green4') # add the line of the best fit into the scatter plot
```


# result
*Conduct a hypothesis test to assess whether there a (linear) relationship between the explanatory and response variable.*
*H0:The loudness is not a significant explanotary variable for predicting energy (β1=0)*
*H1:The loudness is a significant explanotary variable for predicting energy (β1 not equal to 0)*
```{r}
T=(0.05980-0)/0.00561 # use a t-test to determine if loudness is a significant explanotary variable for energy
2*pt(T,df=98,lower.tail = FALSE) # because it's two tails, so we should times two to get the p-value
```
*So p-value=4.506288e-18 which is smaller than 0.05, so we can reject hypothesis and conclude that we have sufficient evidence that the loudness is significant explanatory variable for predicting energy (β1 not equal to 0)*



# Resources 

**Nadin Tamer (2019). Top Spotify Tracks of 2018 [Data file]. Retrieved from https://www.kaggle.com/nadintamer/top-spotify-tracks-of-2018**

**What the dataset is and where it is from.**

**4 Ways that Music and Mathematics are Related. (2019, October 13). Retrieved September 22, 2020, from https://www.omahaschoolofmusicanddance.com/our-blog/4-ways-that-music-and-mathematics-are-related/**

**How music and math related to each other**

**Gaab, N. (2017, May 01). Is There a Link between Music and Math? Retrieved September 22, 2020, from https://www.scientificamerican.com/article/is-there-a-link-between-music-and-math/**

**What is the relationship between math and music.**

**Foundation, C. (n.d.). 12 Foundation. Retrieved September 22, 2020, from https://flexbooks.ck12.org/cbook/ck-12-physics-flexbook-2.0/section/12.3/primary/lesson/intensity-and-loudness-of-sound-ms-ps**

**Use the knowledge of physics to understand the energy and loudness of sound.**

**Malkin, J., Li, X., & Bilmes, J. (2005). Energy and loudness for speed control in the vocal joystick. Retrieved September 22, 2020, from https://ieeexplore.ieee.org/document/1566489/citations?tabFilter=papers**

**It proposes and describes several methods for using speech power as an estimate of intentional loudness, and a mapping from this loudness estimate to a continuous control.**

**Ford, C. (2015, August 26). University of Virginia Library Research Data Services + Sciences. Retrieved September 22, 2020, from https://data.library.virginia.edu/understanding-q-q-plots/**

**It first gives a clear definition of normal QQ plot to help readers quickly understand what normal QQ plot is. More importantly, it combines images to teach readers how to get information from the images of normal QQ plot.**